### 2014-05-17 SPOP/projekt/doc/report/report.tex
Started creating SPOP report

\documentclass[a4paper,12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[MeX]{polski}
\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{pgfplots}
\usepackage{subfigure}
\usepackage{textcomp}

\title{SPOP 2014 L. Projekt. Opis źródeł.}
\author{Aniserowicz Michał, Turek Jakub}
\date{}

\begin{document}

\maketitle

\section{Moduły}

\begin{itemize}
 \item Logic.Game
 \item Logic.AI
 \item Presentation.GameLoop
 \item Presentation.GameSpecific
 \item Presentation.SaveLoad
 \item Main
\end{itemize}

\section{Heurystyka}

\end{document}